---
title: "Statistical Critique 2: Exploring p-values"
author: "Your name here!"
format: html
self-contained: true
editor: visual
---

```{r}
#| label: data-load

library(tidyverse)
library(moderndive)
library(broom)

# Reference the R package you data live inside, since you will need to load that package in!
# Package for your data:

```

# Part One: Revisiting the Midterm Project

For the first part of this critique, you are going to revisit the model you selected for your Midterm Project.

## Visualizations

```{r}
#| label: project-plots
# In the code chunk below, copy-and-paste the plots you created in your Midterm Project.

```

## Model Justification

<!-- Justify (in 2-3 sentences) why you chose this model! -->

# Part Two: Using a p-value Instead

```{r}
#| label: obtain-p-values
# Step 1: fit the most complex regression model
# Step 2: run an ANOVA to obtain p-values

```

**Based on the p-value(s) you obtained from the ANOVA table, what model is the "best" model?**

# Part Three: Learning More about Misuses of $p$-values

**Read the American Statistical Association's statement on $p$-values and 
statistical significance [and]{.underline} the Nature article on 
"statistical significance.**

## Part Four: Reflection

Now that you have compared the model you would have selected using p-values
instead of a visualization and the misuses of p-values, I would like you to 
reflect on the benefits and drawbacks of using p-values. 

1. What are the benefits of using p-values to assess the strength of evidence
in a study? 

2. What are some limitations or drawbacks of relying solely on p-values to
determine whether a result is meaningful or important?

3. What is something you've learned in this activity that you will take with you
in your future courses / research?
