---
title: "Lab 7 Grading Guide"
format: html
editor: visual
---

## Question 1

**To earn a Success**:

-   states the number of rows in the data frame

*If they use `group_by()`*

> Careful! We don't need to make any groups to know how many penguins are in these data! We only need to know the number of rows in the dataset!

## Question 2

**To earn a Success**:

Creates scatterplot with:

-   `latitude` on x-axis
-   `size` on y-axis
-   y-axis label includes units for `size` (millimeters)

::: callout-note
They are not required to include a regression line, but if they do, great!
:::

*If they don't have axis labels with the units*:

> It is important to include an axis label stating what the units of each variable are! If you are wondering what the units of crab size are, you can look up the help file (using ?pie_crab)

## Question 3

**To earn a Success**:

Description includes **all** of the following:

-   form of relationship (linear)
-   direction of relationship (positive)
-   strength of relationship (moderate)
-   location of any outliers

*If they forget one of these*:

> Careful! You were asked to describe the form, direction, strength, and unusual points for the plot. Remember, it is important to explicitly state where you believe the outliers are, so the reader knows where to look! I would also encourage you to justify **why** you believe this is an outlier, since other people might not think so!

*If they don't state [where]{.underline} the outlier is*:

> It is important to explicitly state where you believe the outliers are, so the reader knows where to look! I would also encourage you to justify **why** you believe this is an outlier, since other people might not think so!

## Question 4

**To earn a Success:**

Code should look like the following:

```{r}
#| eval: false
#| echo: true

obs_slope <- obs_slope <- marsh_info %>% 
  specify(response = size, 
          explanatory = latitude) %>% 
  calculate(stat = "slope")
```

## Question 5

**To earn a Success:**

Code should look like the following:

```{r}
#| eval: false
#| echo: true

null_dist <- pie_crab %>% 
  specify(response = size, 
          explanatory = latitude) %>% 
  hypothesize(null = "independence") %>% 
  generate(reps = 500, 
           type = "permute") %>% 
  calculate(stat = "slope")
```

If they chose `reps` larger than 500, that's okay!

*If they chose `reps` smaller than 500*:

> Careful! How many resamples (reps) were you asked to generate?

## Question 6

**To earn a Success:**

Code should look like the following:

```{r}
#| eval: false
#| echo: true

get_p_value(null_dist, 
            obs_stat = obs_slope, 
            direction = "two-sided")

```

*If they don't specify `"two-sided"`*:

> Careful! How many tails are there in a "standard" hypothesis for the slope?

## Question 7

**To earn a Success**: States they reject $H_0$ because the p-value is less than 0.1

*If they do not state whether they reject or fail to reject* $H_0$:

> You were asked to make a decision regarding the hypotheses, which has two possible options. Which option do you choose and why?

*If they say they fail to reject* $H_0$:

> Careful! What p-value did you get? Was it smaller than the alpha I specified? The alpha stipulates how "unusual" the observed statistic needs to be in order for us to reject $H_0$. Is your p-value smaller than this threshold?

## Question 8

**To earn a Success**: States there is a (linear) relationship between crab size and latitude.

> Fill in response!

## Question 9

```{r}
#| eval: false
#| echo: true

bootstrap_dist <- pie_crab %>% 
  specify(response = size, 
          explanatory = latitude) %>% 
  generate(reps = 500, 
           type = "bootstrap") %>% 
  calculate(stat = "slope")
```

If they chose `reps` larger than 500, that's okay!

*If they chose `reps` smaller than 500*:

> Careful! How many resamples (reps) were you asked to generate?

## Question 10

**To earn a Success:**

Code should look like the following:

```{r}
#| eval: false
#| echo: true

get_confidence_interval(bootstrap_dist, 
                        level = 0.90, 
                        type = "percentile")
```

::: callout-note
# `"percentile"` is the default

So, it is okay if they don't specify it!
:::

*If they use a `level` other than 0.90*:

> Look back at the lab instructions, what percentage confidence interval were you asked to construct?

## Question 11

**To earn a Success**:

The interpretation must state:

-   **confidence**: they are 90% confident
-   **statistic**: the slope between crab size and latitude
-   **population**: for all marshes along the eastern US
-   **interval**: is between \[lower bound\] and \[upper bound\]

::: callout-note
Note that every group will get a different interval, due to the randomness of bootstrapping!
:::

*If they don't state their confidence*:

> How much confidence do you have in your interval?

*If they don't state the statistic in context*:

> We need to be specific about the what parameter we believe is in our interval. The slope statistic is measuring the relationship between which variables?

*If they don't include a population:*

> What population does this interval apply to? Where were these crabs sampled from? That is the population your interval applies to!

<!-- *If they have state an incorrect population*: -->

<!-- > Careful! Are you analyzing data on crabs or on marshes? Where were these marshes located? What population would you characterize these 13 marshes belonging to? That is the population your interval applies to! -->

## Question 12

Based on this interval, does Bergmann's Rule apply to Fiddler Crabs? Specifically, does the size of a crab increases as its distance from the equator increases?
